### 2023-12-26 pnrs.tex
yaaay

@@

First, let us remark that $\CorrectPoly \subseteq \CoveredPoly$, because
polynomials in $\CorrectPoly$ are \kl{non-negative}. Let us also check that the
counter example provided in \cref{thm:counter-example} is not in
$\CorrectPoly$. For that, notice that for $P$ defined in
\cref{def:bad-polynomial}, $P(X,Y,1) = 3X^2 + 3Y^2 - 2XY$, which has a negative
coefficient for a \kl{maximal monomial}, namely $-2XY$.

@@

Let us assert that polynomials that have all coefficients in $\Nat$ are
\kl{$\Nat$-rational polynomials}, which will be a base case in an upcoming
induction.

\begin{lemma}
    Let $f$ be a  \kl{$\Nat$-polyregular} function. 
    There exists $\omega \in \Nat$
    such that for all $p \in \Nat$,
    for all $\alpha_0, \dots, \alpha_p \in \Sigma^*$,
    for all $u_1, \dots, u_p \in \Sigma^*$,
    there exists a polynomial $P \in \Rel[X_1, \dots, X_p]$
    whose \kl{maximal monomials} have \kl{positive coefficients},
    and such that:
    \begin{equation*}
        f(\alpha_0 \prod_{i = 1}^p (u_i^{\omega \times X_i} \alpha_i))
        = P(X_1, \dots, X_p) \quad .
    \end{equation*}
\end{lemma}

\begin{lemma}
    The \kl{$\Nat$-rational polynomials} are in $\CorrectPoly$.
\end{lemma}
\begin{proof}[Proof Idea]
    Let $P$ be a \kl{$\Nat$-rational polynomial}.
    Let $\Sigma$  be the finite alphabet such that
    $P \in \Rel[\seqof{X_a}{a \in \Sigma}]$,
    and $f \colon \Sigma^* \to \Nat$ be the \kl{commutative}
    $\Nat$-rational series computing $P$.
    Because $f$ has polynomial growth, and is a $\Nat$-rational series,
    it is a $\Nat$-polyregular function 
    (\cref{polynomial-rational-polyreg:fact}).
    In particular, 
    there exists a finite monoid $M$,
    a morphism $\mu \colon \Sigma^* \to M$,
    an integer $p  \in \Nat$,
    and a production function $\pi \colon M^p \to \Nat$,
    such that for all words $w \in \Sigma^*$,
    \begin{equation*}
        f(w) = \sum_{w = u_1 \cdots u_p} \pi(\mu(u_1), \dots, \mu(u_p)) \quad .
    \end{equation*}
    Because $f$ is \kl{commutative}, we can restrict our attention
    to words of the form $\prod_{a \in \Sigma} a^{k_a}$ to compute $f$.
    To that end, let us
    consider $\omega \in \Nat$ such that
    $m^\omega$ is an idempotent for all $m \in M$ (this is possible because
    $M$ is finite).
    Then, $f( \prod_{a \in \Sigma} a^{ \omega \times k_a} )$
    can be computed as follows:
    \begin{align*}
    f(\prod_{a \in \Sigma} a^{ \omega \times k_a} )
    &= 
    \sum_{\prod_{a \in \Sigma} a^{\omega \times k_a}
    = u_1 \cdots u_p} 
         \pi(\mu(u_1), \dots, \mu(u_p)) \\
    &= \pi(a^\omega, \dots, a^\omega) 
    \end{align*}
    is a polynomial in $k_a$, whose maximal coefficients are
    sums of $\alpha_i$.

    Furthermore, $f(w) = P(\seqof{\card[a]{w}}{a \in \Sigma})$,
    hence we can identify coefficients.
    \textbf{TODO: this has positive maximal coefficients}.
    Furthermore, this remains true when restricting the language.
\end{proof}

@@

To prove the theorem, let us first provide a nice characterization of
\kl{commutative} \kl{$\Rel$-polyregular functions}
in terms of polynomials.

@@

    Furthermore, the polynomials are computable from $f$.
\end{lemma}
\begin{proof}
\end{proof}

@@

>>>>

\usepackage{comment}
%\includecomment{forshort}
%\excludecomment{proof}

@@

First, let us remark that $\CorrectPoly \subseteq \CoveredPoly$, because
polynomials in $\CorrectPoly$ are \kl{non-negative}. Let us also check that the
counter example provided in \cref{thm:counter-example} is not in
$\CorrectPoly$. For that, notice that for $P$ defined in
\cref{def:bad-polynomial}, $P(X,Y,1) = 3X^2 + 3Y^2 - 2XY$, which has a negative
coefficient for a \kl{maximal monomial}, namely $-2XY$. Let us also assert that
polynomials that have all coefficients in $\Nat$ are \kl{$\Nat$-rational
polynomials}, which will be a base case in an upcoming induction.

@@

\begin{lemma}
    \label{n-poly-combinatorics:lem}
    Let $f$ be a  \kl{$\Nat$-polyregular} function. 
    There exists $\omega \in \Nat$
    such that for all $p \in \Nat$,
    for all $\alpha_0, \dots, \alpha_p \in \Sigma^*$,
    for all $u_1, \dots, u_p \in \Sigma^*$,
    there exists a polynomial $P \in \Rel[X_1, \dots, X_p]$
    whose \kl{maximal monomials} have \kl{positive coefficients},
    and such that:
    \begin{equation*}
        f(\alpha_0 \prod_{i = 1}^p (u_i^{\omega \times X_i} \alpha_i))
        = P(X_1, \dots, X_p) \quad .
    \end{equation*}
\end{lemma}
\begin{proof}[Proof Sketch]
    Let $\omega$ be an idempotent power for the finite monoid $M$.
    We define an equivalence between partitions of a word 
    $w \defined \alpha_0 \prod_{i = 1}^p (u_i^{\omega \times X_i} \alpha_i)$
    as follows: two partitions into $k$ words in $\Sigma^*$ are equivalent
    when normalizing them with the rules
    $u_i^{\omega} u_i^{\omega} \to u_i^{\omega}$ for all $1 \leq i \leq p$.
    This equivalence relation has finite index, and all partitions in the same
    equivalence class yield the same value through $\pi$, because
    $\mu(u_i^\omega) = \mu(u_i)^\omega$ is an idempotent element of the monoid $M$.

    Now, let us remark that the number of partitions
    in a certain equivalence class is proportional to
    the number of nanania, that is, is of the form
    $(X_i - K_i)^{n_i}$ for some $K$. We conclude that
    maximal monomials indeed have positive coefficients.
    We even conclude that
    the polynomial is in $\Nat[\vec{X}]$
    if we shift the variables with a large enough $K$. 
\end{proof}

\begin{lemma}
    The \kl{$\Nat$-rational polynomials} are in $\CorrectPoly$.
\end{lemma}
\begin{proof}[Proof Idea]
    Let $P$ be a \kl{$\Nat$-rational polynomial}.
    Let $\Sigma$  be the finite alphabet such that
    $P \in \Rel[\seqof{X_a}{a \in \Sigma}]$,
    and $f \colon \Sigma^* \to \Nat$ be the \kl{commutative}
    $\Nat$-rational series computing $P$.
    Because $f$ has polynomial growth, and is a $\Nat$-rational series,
    it is a $\Nat$-polyregular function 
    (\cref{polynomial-rational-polyreg:fact}).

    Using \cref{n-poly-combinatorics:lem},
    we conclude that maximal monomials of $P$ have
    positive coefficients. Furthermore, when fixing variables
    in $P$, we are computing a new \kl{$\Nat$-polyregular function},
    hence we also conclude that the maximal monomials have
    positive coefficients.
    We have proven that $P \in \CorrectPoly$.
\end{proof}

@@

To prove the theorem, let us first provide a nice characterization of
\kl{commutative} \kl{$\Rel$-polyregular functions} in terms of polynomials.
This is actually a refinement of \cref{n-poly-combinatorics:lem} in the
commutative case.

@@

    Furthermore, the polynomials are computable from $f$.
\end{lemma}
\begin{proof}
    equivalent partitions of the word, but commutativity?.q
\end{proof}

@@

\begin{conjecture}
    star-free $\Nat$-rational is star-free and $\Nat$-rational?
\end{conjecture}
